---
title: "VascularCalcification_BM_TW"
author: "Thijs (Mattheus) Wildschut"
date: "`r Sys.Date()`"
output: html_document
---

Built with R version `r getRversion()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(patchwork)
options(bitmapType='cairo')

```

# Vascular Calcification_Biomarker Analyses {.tabset}

<!-- ## Biomarkers clinical study -->
<!-- To-do: -->
<!-- - Combination of biomarkers? -->
<!-- - Trajectory analyses? -->

<!-- ```{r Biomarkers clinical study, fig.width=18, fig.height=11} -->
<!-- raw.data = data.table::fread("Input/biomarkers.csv") -->
<!-- # unique(paste0(raw.data$`Analysis Visit`, "\n(", raw.data$`Planned Time Point Name`, ")")) -->
<!-- # unique(raw.data$`Biomarker Test Name`) -->

<!-- data = raw.data %>% -->
<!--   mutate(`Biomarker Numeric Result in Std Format` = as.numeric(`Biomarker Numeric Result in Std Format`), -->
<!--          `Analysis Visit` = factor(paste0(`Analysis Visit`, "\n(", `Planned Time Point Name`, ")"), -->
<!--                                    levels = c("Week 1 Day 1\n(Pre-Dose)", "Week 1 Day 1\n(20 Minutes Prior To End of Infusion)", -->
<!--                                               "Week 12 Day 5\n(Pre-Dose)", "Week 24 Day 5\n(Pre-Dose)", "Follow-up\n(NA)"), -->
<!--                                    labels  = c("Week 1 Day 1\n(Pre-Dose)", "Week 1 Day 1\n(20 Minutes Prior\nTo End of Infusion)", -->
<!--                                               "Week 12 Day 5\n(Pre-Dose)", "Week 24 Day 5\n(Pre-Dose)", "Follow-up"))) %>% -->
<!--   group_by(`Unique Subject Identifier`, `Biomarker Test Name`) %>%  -->
<!--   filter("Week 1 Day 1\n(Pre-Dose)" %in% `Analysis Visit`) %>% -->
<!--   # tally() -->
<!--   dplyr::mutate(`Biomarker normalized to baseline` = `Biomarker Numeric Result in Std Format`- -->
<!--                   `Biomarker Numeric Result in Std Format`[`Analysis Visit` == "Week 1 Day 1\n(Pre-Dose)"]) %>% -->
<!--   ungroup() %>% -->
<!--   filter(`Analysis Value Flag` == "Y" & `Biomarker Result in Original Units` != "<LLOQ" & !is.na(`Biomarker normalized to baseline`)) %>% -->
<!--   filter(`Analysis Visit` != "Week 1 Day 1\n(Pre-Dose)")# %>% -->

<!--   # filter(`Biomarker Test Name` == "Iron") -->

<!-- ggplot(data, aes(x = `Analysis Visit`, y = `Biomarker normalized to baseline`, fill = `Description of Planned Arm`)) + -->
<!--   geom_boxplot(alpha = 0.2, outlier.alpha = 0) +  -->
<!--   geom_point(size = 1, position = position_jitterdodge(jitter.width = 0.2), aes(col = `Description of Planned Arm`)) + -->
<!--   theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + -->
<!--   facet_wrap(.~`Biomarker Test Name`, scales = "free_y") + -->
<!--   ggpubr::stat_compare_means(method = "wilcox.test", label = "p.signif") # method = "t.test",  -->

<!-- ``` -->

## SNF472

- Studies: Dog 9 months & Rat 6 months (endpoint measurements only)
- Data: Cmax, Serum Ca, Serum P, and Serum ALP (complete data)
- No bone finding data available (SNF472)

### Correlation plots

```{r SNF472 Correlation plots, fig.width=9.5, fig.height=7}
raw.data2 = readxl::read_excel("Input/20230220 Comparative Toxicologies Bone Parameters_TW.xlsx")
data.472 = raw.data2 %>%
  filter(Compound == "SNF472") %>%
  mutate(Dose = as.factor(paste0(`Dose (mg/kg)`, "mg/kg ", Route, " (", Regime, ")")),
         Dose = factor(Dose, levels = levels(Dose)[order(as.numeric(str_extract(levels(Dose), "[:digit:]+")))]),
         Species = paste0(Species, " (", Duration, ")"),
         Sex = paste0(Sex, " (", N, ")")) %>%
  select(where(~ !all(is.na(.x)))) %>% select(Species, Sex, Dose, `Dose (mg/kg)`, `Cmax SS (µM)`:`Serum ALP (U/L)`) %>%
  mutate(across(`Cmax SS (µM)`:`Serum ALP (U/L)`, as.numeric),
         `Cmax SS (µM)` = replace_na(`Cmax SS (µM)`, 0))

data.472_long = data.472 %>%
  pivot_longer(cols = c(`Dose (mg/kg)`, `Serum Ca (mg/dL)`:`Serum ALP (U/L)`), 
               names_to = "Measurement", values_to = "Value")

plot_list = map(list("Dog (9 months)", "Rat (6 months)"), function(species){
  ggplot(data.472_long %>% filter(Species == species), 
         aes(x = `Cmax SS (µM)`, y = Value, shape = Sex, group = Sex, linetype = Sex)) +
    geom_point(aes(col = Dose), size = 2) +
    scale_color_viridis_d() +
    geom_line() +
    facet_grid(Measurement ~ ., scales = "free_y", switch = "y") +
    labs(y = NULL, title = species) +
    theme_bw() + theme(plot.title = element_text(hjust = 0.5))
})
wrap_plots(plot_list)

```
<hr style="border:1px solid gray"></hr>


### Significance correlations to Cmax
```{r Significance correlations to Cmax, fig.width=7.5, fig.height=3.5}

lm_function = function(title){
  lm_formula = switch(title, "Add 'Sex' to model" = "+ Sex", "")
  sex_filter = switch(title, "Female only" = "Female", "Male only" = "Male", "Female|Male")
  species_list = list("Rat (6 months)", "Dog (9 months)")
  var_list = list("`Dose (mg/kg)`", "`Serum ALP (U/L)`", "`Serum Ca (mg/dL)`", "`Serum P (mg/dL)`")

  pval_df = map_dfr(species_list, function(species){
    map_dfc(var_list, function(var){
        m = lm(formula = as.formula(paste(var, "~ `Cmax SS (µM)`", lm_formula)),
               data = data.472 %>% filter(Species == species & str_detect(Sex, sex_filter)))
        m.sum = summary(m)
        m.sum$coefficients[,"Pr(>|t|)"]["`Cmax SS (µM)`"]
    })
  }) %>% t %>% as.data.frame %>% `colnames<-`(species_list) %>% mutate(Measurement = str_remove_all(unlist(var_list), "`"))
  pval_long = pval_df %>% 
    pivot_longer(-Measurement, names_to = "Species", values_to = "p-value") %>%
    mutate(Title = title)
}

data_plot = map_dfr(list("Female only", "Male only", "Both sexes", "Add 'Sex' to model"), lm_function) %>%
  mutate(Title = factor(Title, levels = c("Female only", "Male only", "Both sexes", "Add 'Sex' to model")))
ggplot(data_plot, aes(x = Species, y = Measurement, fill = -log10(`p-value`), label = signif(`p-value`, 2))) +
  geom_tile(col = "white") + geom_text(col = "white", size = 3.5) +
  facet_grid(.~Title) +
  scale_fill_viridis_c(labels = ~10^-.x, name = "p-value") + scale_y_discrete(limits = rev) +
  labs(x = NULL, y = NULL) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

```

## SNF610
- Studies: Dog 15 days & Rat 15 days (2x studies)
- Data: Cmax, Serum Ca, Serum P, and Serum ALP, one of the 15 day Rats studies measured Serum Mg
  - Complete data, except for one of the 15 day Rats studies in which Serum Mg is measured instead of Serum P
- Bone finding data available: for analysis simplified as "Y" (any kind of finding) or "N" (no finding)
  - For Dog study, all N
  - For Rat studies, comparison of Y vs N possible

### Boxplots bone findings yes/no
```{r Boxplots bone findings yes/no, fig.width = 9, fig.height=3}
data.610 = raw.data2 %>%
  filter(Compound == "SNF610") %>%
  mutate(Dose = as.factor(paste0(`Dose (mg/kg)`, "mg/kg ", Route, " (", Regime, ")")),
         Dose = factor(Dose, levels = levels(Dose)[order(as.numeric(str_extract(levels(Dose), "[:digit:]+")))]),
         Species = paste0(Species, " (", Duration, ")"),
         Sex = paste0(Sex, " (", N, ")")) %>%
  select(where(~ !all(is.na(.x)))) %>% select(Species, Sex, Dose, `Dose (mg/kg)`, `Cmax SS (µM)`:`Serum Mg (mg/dL)`, `Findings (Y/N)`) %>%
  mutate(across(`Cmax SS (µM)`:`Serum Mg (mg/dL)`, as.numeric),
         `Cmax SS (µM)` = replace_na(`Cmax SS (µM)`, 0))
data.610_long = data.610 %>%
  pivot_longer(cols = c(`Dose (mg/kg)`, `Serum Ca (mg/dL)`:`Serum Mg (mg/dL)`), 
               names_to = "Measurement", values_to = "Value")

data_plot = data.610_long %>% filter(Species == "Rat (15 days)")
ggplot(data_plot, aes(x = `Findings (Y/N)`, y = Value, fill = `Findings (Y/N)`, col = `Findings (Y/N)`)) +
  geom_boxplot(alpha = 0.25, outlier.alpha = 0) + geom_jitter(alpha = 0.5, width = 0.2) +
  ggpubr::stat_compare_means(comparisons = list(c(1,2)), size = 3.5) +
  facet_wrap(.~Measurement, scales = "free_y", nrow = 1) + 
  labs(x = NULL, y = NULL, title = "15-day rat studies\n(data 2x studies combined)") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  theme_bw() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(hjust = 0.5))

```

## INS-3001

- Studies: Dog 7 days, 4-week, and 16-week & Rat 14 days, 4-week, and 13-week
  - For some studies, longitudinal measurements are available
- Data: Cmax, Serum Ca, Serum P, and Serum ALP
  - Not complete data: variation study to study, also sometimes missing within study/timepoint
- Some bone finding data available, perhaps possible to stratify by dose?

```{r, fig.width=9, fig.height=6}
data2 = raw.data2 %>%
  filter(!is.na(N)) %>%
  mutate(Species2 = paste0(Species, "-", Sex, "(N=", N, ")"),
         `Urine Ca (mg/dL)` = as.numeric(`Urine Ca (mg/dL)`),
         Dose2 = as.factor(`Dose (mg/kg)`),
         Duration = factor(Duration, levels = unique(raw.data2$Duration)[c(2,1,6,3,4,7,8,5)]),
         Dur_num = str_remove(Duration, " days|-week| months"),
         Day = ifelse(!is.na(Day), Day, 
                      ifelse(str_detect(Duration, " days"), as.numeric(str_remove(Duration, " days")),
                             ifelse(str_detect(Duration, "-week"), as.numeric(str_remove(Duration, "-week"))*7,
                                    ifelse(str_detect(Duration, " months"), as.numeric(str_remove(Duration, " months"))*30, NA)))),
         Day2 = as.factor(Day)) %>%
  filter(Compound == "INS-3001")
data.sum2 = data2 %>%
  group_by(Species, Sex, Day2, Dose2) %>% 
  summarise(N = sum(N),
            across(is.numeric, ~mean(.x, na.rm = TRUE)))
plot_function = function(fill_var){
  ggplot(data.sum2, aes(x = Day2, y = Dose2, fill = .data[[fill_var]], col = Sex, label = N)) +
    geom_tile(lwd = 0.5, width = 0.4, position = position_nudge(x = ifelse(data.sum2$Sex == "Male", 0.22, -0.22))) +
    geom_text(position = position_nudge(x = ifelse(data.sum2$Sex == "Male", 0.21, -0.21)), size = 3) +
    facet_grid(.~Species, scales = "free", space = "free") +
    scale_fill_viridis_c(na.value = "white") +
    labs(x = NULL, y = "Dose (mg/kg daily s.c.)") +
    theme_bw() + theme(axis.text.x = element_text(angle = 90), panel.grid = element_blank(), panel.grid.major.x = element_line(color = "grey"))
}
# (plot_function("Urine Ca (mg/dL)") + plot_function("Urine Ca (mg/dL)")) /
#   (plot_function("Serum Ca (mg/dL)") + plot_function("Serum Ca (mg/dL)")) +
#   plot_layout(guides = "collect")
# plot_function("Dog", "Urine Ca (mg/dL)") + plot_function("Rat", "Urine Ca (mg/dL)") + plot_layout(guides = "collect")

# ggsave("Output/CaLevels_INS-3001.pdf", width = 6.5, height = 7)

```

### Line plot dose & biomarkers over time
```{r, fig.width=7, fig.height=4.5}
# plot_function = function(fill_var){
#   ggplot(data.sum2, aes(x = Day, y = `Dose (mg/kg)`, fill = .data[[fill_var]], col = Sex, label = N)) +
#     geom_tile(lwd = 0.5, width = 0.4, position = position_nudge(x = ifelse(data.sum2$Sex == "Male", 0.22, -0.22))) +
#     geom_text(position = position_nudge(x = ifelse(data.sum2$Sex == "Male", 0.21, -0.21)), size = 3) +
#     facet_grid(.~Species, scales = "free", space = "free") +
#     scale_fill_viridis_c(na.value = "white") +
#     labs(x = NULL, y = "Dose (mg/kg daily s.c.)") +
#     theme_bw() + theme(axis.text.x = element_text(angle = 90), panel.grid = element_blank(), panel.grid.major.x = element_line(color = "grey"))
# }
# plot_function("Urine Ca (mg/dL)") / plot_function("Serum Ca (mg/dL)")

data.sum2 = data2 %>%
  group_by(Species, Sex, Duration, Day2, Dose2) %>% 
  summarise(N = sum(N),
            across(is.numeric, ~mean(.x, na.rm = TRUE)))

data.plot = data.sum2 %>%
  mutate(Dose3 = ifelse(`Dose (mg/kg)` == 30.5, 30, ifelse(`Dose (mg/kg)` == 104, 100, ifelse(`Dose (mg/kg)` == 304, 300, `Dose (mg/kg)`))))
data.plot = data.plot %>%
  mutate(Dose3 = factor(Dose3, levels = sort(unique(data.plot$Dose3))))

# plotly::ggplotly(
  ggplot(data.plot, aes(x = Day, y = `Urine Ca (mg/dL)`, col = Dose2, label = N)) + #interaction(Dose2, Duration)
    # geom_hline(yintercept = min(data.plot$`Urine Ca (mg/dL)`, na.rm = TRUE), linetype = 2, col = "grey") +
    geom_line(aes(grouping = Dose2)) + geom_point(alpha = 0.8, aes(text = Duration, size = N)) +
    # geom_text(position = position_nudge(x = ifelse(data.sum2$Sex == "Male", 0.21, -0.21)), size = 3) +
    facet_grid(Species~Sex, scales = "free") +
    scale_color_viridis_d(name = "Dose\n(mg/kg daily s.c.)") +
    guides(color = guide_legend(ncol = 2)) +
    scale_size_continuous(range = c(1,3)) +
  # labs(color = "Dose\n(mg/kg daily s.c.)") +
    theme_bw() #, #+ theme(axis.text.x = element_text(angle = 90), panel.grid = element_blank(), panel.grid.major.x = element_line(color = "grey"))
# )

```

