---
title: "RESOLUTE_Mitoferrin-2_Analyses"
author: "Thijs (Mattheus) Wildschut"
date: "`r Sys.Date()`"
output: html_document
---

Built with R version `r getRversion()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
setwd("C:/Users/U0142615/OneDrive - CSL Behring/Documents/R/Projects/R_CSL-Vifor_TW/RESOLUTE")
# source("../SourceFile_TW.R")
```

```{r load packages & functions, include=FALSE}

```

# RESOLUTE_Mitoferrin-2_Analyses {.tabset}

## 1B-C: Ionomics
```{r Mitoferrin-2 Ionomics, fig.width=5.5, fig.height=4.5}
bca = fread("Input/BCA_TW.csv") %>%
  filter(Fraction == "Cytosolic")
raw.data = fread("Input/Ionomics_MIU03_TW.csv") 
data = raw.data %>%
  dplyr::mutate(across(-Name & -Condition, as.numeric)) %>%
  left_join(bca, by = "Name") %>%
  dplyr::mutate(across(contains(c("(MR)", "(HR)")), ~.x/Concentration*1000)) %>%
  `colnames<-`(str_remove(colnames(.), "\\(.*\\)")) %>%
  dplyr::select(-Code & -Concentration & -Fraction) %>%
  pivot_longer(cols = -Name & -Condition, names_to = "Ion", values_to = "Concentration") %>%
  mutate(Condition = factor(Condition, levels = c("DOX-", "DOX+")),
         Ion = factor(Ion, levels =  c("Fe58", "Fe56", "Mn55", "Cu63", "Zn66"))) %>%
  filter(!is.na(Ion)) %>%
  group_by(Ion, Condition) %>% add_tally %>%
  dplyr::mutate(Mean = mean(Concentration, na.rm = TRUE), sd = sd(Concentration, na.rm = TRUE),
                se = sd/sqrt(n), Max = max(Concentration, na.rm = TRUE))
data.fold = data %>%
  group_by(Ion) %>%
  dplyr::summarise(FC = mean(Concentration[Condition == "DOX+"])/mean(Concentration[Condition == "DOX-"]))
  

ggplot(data, aes(x = Condition, y = Concentration, group = Condition, col = Condition)) +
  stat_summary(fun.y = "mean", fun.ymin = "mean", fun.ymax= "mean", width = 0.4, geom = "crossbar", col = "black") +
  geom_point() +
  scale_color_manual(values = c("DOX+" = "grey40", "DOX-" = "grey80")) +
  facet_wrap(.~Ion, scales = "free", nrow = 2) +
  # geom_errorbar(aes(ymin=Mean, ymax=Max), width=.2) +
  labs(x = NULL, y = "Ion concentration (ug/L)\nNormalized to protein (mg/ml)") +
  theme_bw() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0,0.125))) +
  stat_compare_means(comparisons = list(c(1,2)), method = "t.test", method.args = list(var.equal = TRUE), size = 4)

# ggsave("Output/Mitof2_Ionomics_Dots.pdf", width = 5.5, height = 4.5)

```

## RPA assay

### Load RPA data
```{r Load RPA data}
RPA.data = fread("Input/RPA-data_TW.csv", stringsAsFactors = TRUE) %>%
  as.data.frame %>%
  mutate(Software = factor(Software, levels = c("ScanR", "Columbus Cytoplasm", "Columbus Find Spots")),
         Concentration = factor(Concentration, levels = c("No Fe2+ added", "10uM Fe2+", "100uM Fe2+")),
         Condition = factor(Condition, levels = c("DOX-", "DOX+")),
         Celltype = factor(Celltype, levels = c("HEKiM2", "HEK-JI-WT")))

RPA.plot = function(data){
  ggplot(data, aes(x = Condition, y = RPA, fill = Condition, col = Condition)) +
    geom_boxplot(alpha = 0.25, outlier.alpha = 0) + geom_jitter(alpha = 0.75, width = 0.25) +
    xlab(NULL) + ylab("RPA intensity (AU)\nAverage per well") +
    theme_bw() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    stat_compare_means(comparisons = list(c(1,2)), method = "t.test", method.args = list(var.equal = TRUE),
                       label = "p.signif", vjust = 0.4, hide.ns = TRUE) +
    scale_y_continuous(expand = expansion(mult = c(0,0.1)), limits = c(0,NA)) +
    scale_color_manual(values = c("DOX+" = "grey40", "DOX-" = "grey80")) + 
    scale_fill_manual(values = c("DOX+" = "grey40", "DOX-" = "grey80"))
}
```

### 2B: HEKiM2 vs HEK-JI-WT comparison

```{r HEK-JI-WT comparison, fig.width=4, fig.height=2.4}
data = RPA.data %>%
  filter(Plot == "HEK-JI-WT")

RPA.plot(data) +
  facet_grid(.~Celltype, scales = "free")
# ggsave("Output/RPA_HEK-JI-WT.pdf", width = 4, height = 2.4)

```

### 2C: Time course (0-50 minutes)
```{r RPA Time course, fig.width=4.5, fig.height=3}
data = RPA.data %>%
  filter(Plot == "Timecourse")

data.sum = data %>%
  group_by(Condition, Timepoint) %>%
  dplyr::mutate(N = length(RPA), Mean = mean(RPA), sd = sd(RPA), se = sd/sqrt(N))

ggplot(data.sum, aes(x = Timepoint, y = Mean, fill = Condition, col = Condition, group = Condition)) +
  # geom_point(aes(y = RPA))
  geom_errorbar(aes(ymin = Mean-se, ymax = Mean+se), width = 2.5) +
  geom_line() + geom_point() +
  scale_y_continuous(expand = expansion(mult = c(0,0.05)), limits = c(0,NA)) +
  scale_color_manual(values = c("DOX+" = "grey40", "DOX-" = "grey80")) +
  xlab("Time point\nMinutes after RPA addition") + ylab("RPA intensity (AU)\nMean per well") +
  theme_bw()

# ggsave("Output/RPA_Timecourse.pdf", width = 4.5, height = 3)

```

### 2E: Software comparison

```{r Software comparison, fig.width=5.5, fig.height=5.25}
data = RPA.data %>%
  filter(Plot == "Software") %>%
  group_by(Software, Concentration) %>% 
  dplyr::mutate(FC = mean(RPA[Condition == "DOX-"])/mean(RPA[Condition == "DOX+"]))

RPA.plot(data) +
  facet_grid(Software~Concentration, scales = "free")

# ggsave("Output/RPA_Software.pdf", width = 5.5, height = 5.25)

```

### 2F: Serum comparison

```{r Serum comparison, fig.width=5.5, fig.height=4}
data = RPA.data %>%
  filter(Plot == "Serum")

RPA.plot(data) +
  facet_grid(Serum~Concentration, scales = "free")

# ggsave("Output/RPA_Serum.pdf", width = 5.5, height = 4)

```

### 2F: Serum comparison z' calculation

```{r Serum comparison, fig.width=5.5, fig.height=4}
data = RPA.data %>%
  filter(Plot == "Serum")

data2 = data %>%
  group_by(Serum, Condition, Concentration) %>% dplyr::summarise(Mean = mean(RPA), SD = sd(RPA)) %>%
  group_by(Serum, Concentration) %>% 
    dplyr::summarise(Zprime = 1-(3*(SD[Condition == "DOX+"]+SD[Condition == "DOX-"]))/
                                      abs(Mean[Condition == "DOX+"]-Mean[Condition == "DOX-"]))#,
  #                                   SD3 = 3*(SD[Condition == "DOX+"]+SD[Condition == "DOX-"]),
  #                                   MeanDiff = Mean[Condition == "DOX+"]-Mean[Condition == "DOX-"])
```

### 2G: Passage variability

```{r Passages comparison, fig.width=5.5, fig.height=2.4}
data = RPA.data %>%
  filter(Plot == "Passages")

RPA.plot(data) +
  facet_grid(Serum~Passage, scales = "free")

# ggsave("Output/RPA_Passages.pdf", width = 5.5, height = 2.4)

```

### 2G: Passage variability z' calculation

```{r Passages comparison, fig.width=5.5, fig.height=2.4}
data = RPA.data %>%
  filter(Plot == "Passages")

data2 = data %>%
  group_by(Passage, Condition) %>% dplyr::summarise(Mean = mean(RPA), SD = sd(RPA)) %>%
  group_by(Passage) %>% dplyr::mutate(Zprime = 1-(3*(SD[Condition == "DOX+"]+SD[Condition == "DOX-"]))/
                                        abs(Mean[Condition == "DOX+"]-Mean[Condition == "DOX-"]),
                                      SD3 = 3*(SD[Condition == "DOX+"]+SD[Condition == "DOX-"]),
                                      MeanDiff = Mean[Condition == "DOX+"]-Mean[Condition == "DOX-"])

# ggsave("Output/RPA_Passages.pdf", width = 5.5, height = 2.4)

```


### One function to plot them all
```{r RPA function}
RPA.data = fread("Input/RPA-data_TW.csv", stringsAsFactors = TRUE) %>%
  as.data.frame %>%
  mutate(Software = factor(Software, levels = c("ScanR", "Columbus Cytoplasm", "Columbus Find Spots")),
         Concentration = factor(Concentration, levels = c("No Fe2+ added", "10uM Fe2+", "100uM Fe2+")),
         Condition = factor(Condition, levels = c("DOX-", "DOX+")))

RPA.function = function(variable, facet1, facet2){
  data = RPA.data %>%
    filter(Plot == variable) %>%
    mutate(Facet = .[,facet2])
  
  ggplot(data, aes(x = Condition, y = RPA, fill = Condition, col = Condition)) +
    geom_boxplot(alpha = 0.25, outlier.alpha = 0) + geom_jitter(alpha = 0.75, width = 0.25) +
    facet_grid(Facet ~ .data[[facet1]], scales = "free") +
    xlab(NULL) + ylab("RPA intensity (AU)\nAverage per well") +
    theme_bw() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    stat_compare_means(comparisons = list(c(1,2)), method = "t.test", method.args = list(var.equal = TRUE),
                       label = "p.signif", vjust = 0.4, hide.ns = TRUE) +
    scale_y_continuous(expand = expansion(mult = c(0,0.1)), limits = c(0,NA)) +
    scale_color_manual(values = c("DOX+" = "grey40", "DOX-" = "grey80")) + 
    scale_fill_manual(values = c("DOX+" = "grey40", "DOX-" = "grey80"))
}
RPA.function("HEK-JI-WT", "Celltype", "Concentration")
RPA.function("Software", "Concentration", "Software")
RPA.function("Software", "Concentration", "Software")


```

## RESOLUTE transcriptomics
Data exported from RESOLUTE transcriptomics dashboard \
Comparison WT vs OE (HEK): https://re-solute.eu/internal/database/dashboards/transcriptomics/analysis/AN0298-B \
Comparison KO vs OE (HCT116): https://re-solute.eu/internal/database/dashboards/transcriptomics/analysis/AN010S-7 \

### Overlap significant changes
```{r Overlap significant changes, fig.width=10, fig.height=4}
data.WT = read.csv("Input/RESOLUTE_SCL25A28_WT-OE.csv")# %>% mutate(Comparison = "WT-OE")
data.KO = read.csv("Input/RESOLUTE_SCL25A28_KO-OE.csv")# %>% mutate(Comparison = "KO-OE")
data = merge(data.WT, data.KO, by = c("Gene.Symbol", "HGNC.ID", "Gene.Name", "Ensembl.Gene.ID"), suffixes = c("_WT", "_KO")) %>%
  mutate(SignedP_WT = -log10(adjusted.p.value_WT+1e-50)*sign(log2.fold.change_WT),
         SignedP_KO = -log10(adjusted.p.value_KO+1e-50)*sign(log2.fold.change_KO),
         Direction = factor((SignedP_WT > -log10(0.05) & SignedP_KO > -log10(0.05)) | SignedP_WT < log10(0.05) & SignedP_KO < log10(0.05),
                              levels = c("TRUE", "FALSE"), labels = c("Same", "Opposite")),
         Significant = adjusted.p.value_WT < 0.05 & adjusted.p.value_KO < 0.05)
# plot.log2FC = ggplot(data, aes(x = log2.fold.change_WT, y = log2.fold.change_KO)) +
#   geom_point() +
#   geom_label_repel(data = filter(data, log2.fold.change_WT > 0.1 & log2.fold.change_KO > 0.1), aes(label = Gene.Symbol), max.overlaps = Inf)
# plot.signedP = 
plot = ggplot(data, aes(x = SignedP_WT, y = SignedP_KO)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) +
  geom_point(aes(col = Significant)) +
  scale_color_manual(values = c("TRUE" = "black", "FALSE" = "grey70"), breaks = c("TRUE", "FALSE")) +
  scale_fill_manual(values = c("Same" = "green3", "Opposite" = "red2")) +
  geom_label_repel(data = filter(data, Significant), aes(label = Gene.Symbol, fill = Direction),
                   max.overlaps = Inf, size = 3, label.padding = 0.1, alpha = 0.8) +
  theme_bw()
plot2 = plot + coord_cartesian(xlim = c(-15,15), ylim = c(-15,15))
plot + plot2
# ggsave("Output/SLC25A28-Transcriptomics_WT-OE_KO-OE.pdf", width = 10, height = 4)

```
<hr style="border:1px solid gray"></hr>

### Transcriptomics GSEA & overlap
```{r Transcriptomics GSEA, fig.width=9, fig.height=4.5}
kegg2 = getKEGGPathwayNames(species="hsa")
kegg1 = getGeneKEGGLinks(species="hsa") %>%
  mutate(Gene = AnnotationDbi::mapIds(org.Hs.eg.db::org.Hs.eg.db, GeneID, column = "SYMBOL", keytype = "ENTREZID"),
         PathwayID = str_remove(PathwayID, "path:"),
         Description = str_remove(kegg2$Description[match(PathwayID, kegg2$PathwayID)], " - Homo sapiens \\(human\\)")) %>%
    filter(!str_detect(PathwayID, "hsa05|04930|04940|04950|04936|04932|04931|04933|04934|01501|01502|01503|01521|01524|01523|01522"))  %>%
  select(Description, Gene)

list.kegg = sort(setNames(data$SignedP_KO, data$Gene.Symbol), decreasing = TRUE)
set.seed(seed = 1)
enr.KEGG = GSEA(geneList = list.kegg, TERM2GENE = kegg1, eps = 0, pvalueCutoff = 1)

# enr.GO.up = enrichGO(data$Gene.Symbol[data$SignedP_WT > -log10(0.05)], OrgDb = "org.Hs.eg.db", ont = "BP", keyType = "SYMBOL",
#                   universe = data$Gene.Symbol)@result
# enr.GO.down = enrichGO(data$Gene.Symbol[data$SignedP_WT < log10(0.05)], OrgDb = "org.Hs.eg.db", ont = "BP", keyType = "SYMBOL",
#                   universe = data$Gene.Symbol)@result

data.plot = rbind.data.frame(slice_max(enr.KEGG@result, NES, n = 10), slice_min(enr.KEGG@result, NES, n = 10)) %>%
  mutate(Description = factor(Description, levels = Description),
         Sign = factor(sign(NES), levels = c(-1,1), labels = c("Negative", "Positive")),
         Significant = `p.adjust` < 0.05)

ggplot(data.plot, aes(x = NES, y = Description, size = -log10(`p.adjust`), col = setSize, alpha = Significant)) +
  geom_point() + scale_color_viridis() + scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.2)) + guides(alpha = FALSE) +
  facet_wrap(vars(Sign), scales = "free") +
  labs(x = "NES (Normalized enrichment score)", y = NULL,
       title = "Top 10 enriched KEGG terms\nfrom gene set enrichment analysis (GSEA)\non ADPDK-group Spearman correlations") +
  scale_y_discrete(limits = rev) + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
# ggsave("Output//GSEA_Spearman_ADPKD-Group_Correlations.pdf", width = 12, height = 4.5)

```
<hr style="border:1px solid gray"></hr>

### Transcriptomics Friedreich's ataxia
```{r GSE102008, fig.width=9, fig.height=4.5}
gse = getGEO("GSE102008", GSEMatrix = TRUE)[[1]]
gene.annot = gse@featureData@data
pat.annot = gse@phenoData@data

data = gse@assayData$exprs %>% as.data.frame %>%
  mutate(Gene = gene.annot$ILMN_Gene[match(rownames(.), gene.annot$ID)]) %>%
  group_by(Gene) %>% dplyr::summarise(across(everything(), max))
data2 = data %>% column_to_rownames("Gene") %>%
  t %>% as.data.frame %>%
  mutate(Disease = factor(pat.annot$`status:ch1`[match(rownames(.), pat.annot$geo_accession)],
                          levels = c("Control", "Carrier", "Patient")))

ggplot(data2, aes(x = Disease, y = SLC25A28)) +
  geom_boxplot() +
  stat_compare_means(comparisons = list(c(1,2), c(1,3), c(2,3)), method = "t.test")

```
